---
title: "reshaping_pllr_td60"
output: html_document
---

```{r}
library(dplyr)
library(stringr)
library(readr)
library(tidyr)
```


```{r}
path_td60 <- "../../data/raw/td60_w2tg_pllr/td60phonewise_pllr_df_max.csv"
file.info(path_td60)$size #bytes
file.info(path_td60)$size / 1024^2 #MB
file.info(path_td60)$size / 1024^3 #GB

n_lines <- length(readLines(path_td60, warn = FALSE))
n_lines
```
```{r}
df_pllr_td60<-read_csv(path_td60, show_col_types = FALSE)
glimpse(df_pllr_td60)
```
```{r}
unique(df_pllr_td60$FileName)

```



# The columns in the old data csv file we were using are:
  speaker = col_character(),
  age_months = col_double(),
  filename = col_character(),
  tocs_item = col_character(),
  utterance = col_character(),
  expected_phoneme = col_character(),
  most_likely_phoneme = col_character(),
  position = col_double(),
  phoneme = col_character(),
  prob = col_double()

Ask to Tristan: how to get columns utterance and position

```{r}

df_pllr_td60_long <- df_pllr_td60%>%
  mutate(
    age_months = as.integer(str_match(FileName, "^[^_]*_([^_]*)_")[,2]),
     tocs_item  = str_match(FileName, "-(.+)$")[,2]
         )%>%
  tidyr::pivot_longer(
    cols = starts_with("Phoneme_"),
    names_to = "phoneme",
    values_to = "prob"
  ) %>%
  mutate(phoneme = str_remove(phoneme, "^Phoneme_")) %>%
  filter(!phoneme %in% c("[UNK]", "[PAD]")) %>%
  dplyr::rename(
    speaker = Speaker,
    filename = FileName,
    expected_phoneme = ExpectedPhoneme,
    most_likely_phoneme = MostLikelyPhoneme
  )%>%select(
    speaker, age_months, filename, tocs_item, expected_phoneme, most_likely_phoneme, phoneme,prob
  )
```

```{r}
df_pllr_td60_long|> write_csv("../../data/raw/td60pllr/long_tables/td60phonewise_pllr_df_max_long.csv")
```

### repetir para el csv de pllr adultos:
(este es el archivo que viene de la carpeta ASU)
```{r}
path_adults <- "../../data/raw/adults_w2tg_pllr/td60phonewise_pllr_df_max.csv"
file.info(path_adults)$size #bytes
file.info(path_adults)$size / 1024^2 #MB
file.info(path_adults)$size / 1024^3 #GB

n_lines <- length(readLines(path_adults, warn = FALSE))
n_lines
```
```{r}
df_pllr_adults<-read_csv(path_adults, show_col_types = FALSE)
glimpse(df_pllr_adults)
```

```{r}
unique(df_pllr_adults$FileName)

```

```{r}

df_pllr_adults_long <- df_pllr_adults %>%
  mutate(
    age_code = stringr::str_match(FileName, "^[^_]*_([^_]*)_")[,2],
    age_years = as.integer(substr(age_code, 1, 2)),
    age_extra_months = as.integer(substr(age_code, 3, 4)),
    age_months = 12 * age_years + age_extra_months,
    tocs_item = stringr::str_match(FileName, "-(.+)$")[,2]
  )%>%
  tidyr::pivot_longer(
    cols = starts_with("Phoneme_"),
    names_to = "phoneme",
    values_to = "prob"
  ) %>%
  mutate(phoneme = str_remove(phoneme, "^Phoneme_")) %>%
  filter(!phoneme %in% c("[UNK]", "[PAD]")) %>%
  dplyr::rename(
    speaker = Speaker,
    filename = FileName,
    expected_phoneme = ExpectedPhoneme,
    most_likely_phoneme = MostLikelyPhoneme
  )%>%select(
    speaker, age_months, filename, tocs_item, expected_phoneme, most_likely_phoneme, phoneme,prob
  )
```

```{r}
df_pllr_adults_long|> write_csv("../../data/raw/adults_w2tg_pllr/adults_phonewise_pllr_df_max_long.csv")
```


```{r}
glimpse(df_pllr_adults_long)

```

